---
title: "Packages for EDA"
author: "KPI"
date: "2024_10_04"
output:
  word_document: default
  'officedown:': rdocx_document
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
always_allow_html: true
params:
  directory:
    value: /Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/
---

#### Setup chunk

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
  include = TRUE,
  message = FALSE,
  results = TRUE,
	warning = FALSE
  )
Knit_to_Word = TRUE
```

#### Load packages, options, caption functions, and source file

```{r load packages, include=FALSE}
suppressPackageStartupMessages({
library(beanplot)
library(blandr)
library(BlandAltmanLeh)
library(captioner)
library(car)
library(corrplot)
library(cowplot)
library(CGPfunctions)
library(DataExplorer)
library(data.table)
library(dplyr)
library(EnvStats)
library(flextable)
library(formattable)
library(GGally)
library(ggExtra)  
library(ggpirate)
library(ggplot2)
library(ggpubr) 
library(ggthemes)
library(glue)
library(Hmisc)
library(inspectdf)
library(janitor)
library(kableExtra)
library(likelihoodR)
library(MANOVA.RM)
library(mgcViz)
library(officedown)
library(RColorBrewer)
library(readr)
library(readxl)
library(scales)
library(skimr)
library(SmartEDA)
library(summarytools)
library(tidyquant)
library(tidyr)
library(tidyverse)
library(webshot2)
library(XICOR)
library(zeallot)
})

set_flextable_defaults(font.family = "Times New Roman",
                       border.color = "gray",
                       font.size = 10,
                       digits = 3,
                       theme_fun = "theme_zebra")

options(scipen = 999) 
options(digits = 2)
options(prType='html')

source(glue('{params$directory}','Lab_Notebooks/library.R'))

```

#### Creating table and figure captions

```{r Captions,results=FALSE, echo=TRUE}
fig_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")

table_nums(name = "samples", caption = "Small sample table that is a subset of the AllResults table")
fig_nums(name = "scatter", caption = "Scatterplot of stutter present vs. stutter absent for all samples and both programs in one plot")


```

### Introduction and purpose

From this website (https://www.business-science.io/code-tools/2024/10/03/top-10-r-packages-for-eda.html)

Several packages for doing a quick but comprehensive EDA overview are given. This simply repeats the one's I find of interest. 


#### Creating data tables/frames for use in analysis

The AllResults_Table10.csv contains the most current data, consisting of 38 variables and 8000 rows

"Sample" "RepNum" "RepSetID"  "ConditionID" "TotlogLR"    "Eff"         "Program"     "Stutter"     "NumCont"     "TCon"        "TotDNA"      "amtDNA"      "APH"         "SampleFr"    "Calc_Prop"   "C1"          "C2"          "C3"          "C4"          "Dose1"       "Dose2"       "Dose3"  "Dose4"   "CSF1PO"      "D13S317"     "D16S539"     "D18S51"      "D19S433"     "D21S11"      "D2S1338"     "D3S1358"     "D5S818"      "D7S820"     
"D8S1179"     "FGA"         "TH01"        "TPOX"        "VWA" 


```{r}
MT <- readRDS("/Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/Data/Output_files/AllResults_Data/AllResults_Table_10.rds")

KPIFlexTable(head(MT), height = "100%", dimnames(MT)[[2]])
```

Further exploration (after running skim) showed that the column types are not preserved (duh, it's a csv file). Searching stackoverflow shows the answer is to convert the required columns to factors once, then use saveRDS to (which is an R type of file) to preserve the column type. Also documented in my thesis Dissertation Notes in OneNote

```{r eval=FALSE, include=FALSE}
#Call the function as change_column_type(dt, cols, type = *in quotes as in function*)
cols <- c("NumCont", "C1", "C2", "C3", "C4", "Dose1", "Dose2", "Dose3", "Dose4", "Program", "Stutter", "TCon") 
MT <- change_column_type(MT, cols, type = c("as.factor"))

saveRDS(MT, file = "/Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/Data/Output_files/AllResults_Data/AllResults_Table_10.rds")
```

This was done once, and will not be repeated. 

skimr provides a convenient and elegant summary of the data.
```{r}

skim(MT)

```

skim now shows 12 columns as factors and 26 as numeric. 

```{r results="asis"}
dfSummary(MT, plain.ascii  = FALSE, 
          style        = "grid", 
          graph.magnif = 0.75, 
          valid.col    = FALSE,
          tmp.img.dir  = "/tmp")
```
This gives a nice html summary, but a sucky, unusable print summary. 
```{r}
corrplot(
  cor(MT[,.(SampleFr, Calc_Prop, APH)]),
      method = "circle", 
      addCoef.col = 'grey', 
      order = 'alphabet',
      rect.col = 'blue',
      addrect = 2)
        

```

This doesnt' work on my data

```{r}
chart.Correlation(
  MT[,.(TotDNA,amtDNA,SampleFr, Calc_Prop, APH)],
  histogram = TRUE,
  pch = 19
)
```


```{r}

ggpairs(
  data = MT,
  columns = c("TotDNA","amtDNA","SampleFr", "Calc_Prop", "APH"),
  mapping = aes(color = Stutter)
)
```

```{r}


create_report(
  MT,
  output_dir = "/Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/Lab_Notebooks/Ch_EDA_Examples/tmp",
  output_file = "MasterTable_report.docx"
)
```

This one is probably the best summary of statistical stuff. 

As expected, this notebook will not print to Word. Most are sent to html, others inline. While useful, probably won't make it into the thesis. 

```{r eval=FALSE, include=FALSE}
# ifelse statement to toggle between kableExtra table when knitting to html or flextable when knitting to Word
`
if (Knit_to_Word) {
  KPIFlexTable(DATAFRAME, COLUMNS_TO_PRINT) # Frequently columns to print are of the form 'dimnames(Partial)[[2]][1:14]'
  } else {
  KPITable(DATAFRAME) 
}

```

### Analysis

#### [**Appendix**]{.underline}

#### Using the captioner R package

Inserting a caption:

1.  A function exists for both tables and figures [e.g., table_nums \<-
    captioner(prefix = "Table")] at the beginning of the notebook.

2.  Use the function to create captions as needed. The captions are all
    placed in one chunk at the top of the notebook, and then referenced
    when needed throughout the notebook. [e.g., table_nums(name =
    "samples", caption = "Small sample table that is a subset of the
    AllResults table")]

    1.  The two primary parts of the function is the "name =" which is
        the short name of the caption, and the actual prose to use in
        the caption "caption =" as in the example above.

3.  When the table or figure caption is to be used, it is first placed
    outside of a code **chunk** in a **code block**, as follows (using
    backticks, not apostrophes):

`r table_nums("samples")`

#### Other editing tricks when using an R Notebook

To indent a line in markdown, use '&nbsp'. Eventually RStudio converts
it to a colored dot, but it works.

To insert space between paragraphs, use, "Insert \| Paragraph" in Visual
mode

```{r get-labels, echo = FALSE}
labs = knitr::all_labels()
labs = setdiff(labs, c("setup", "get-labels", "load packages"))
```

```{r all-code, include = !Knit_to_Word, ref.label=labs, eval=FALSE, echo=TRUE}

```

### Reproducibility Section

```{r Reproducibility-SessionInfo-R-environment, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center", out.width='100%', results='asis'}
library("devtools")

if (Knit_to_Word) {
  
  } else {

df_session_platform <- devtools::session_info()$platform %>% 
  unlist(.) %>% 
  as.data.frame(.) %>% 
  rownames_to_column(.)

colnames(df_session_platform) <- c("Setting", "Value")

kable(
  df_session_platform, 
  booktabs = T, 
  align = "l",
  caption = "Reproducibility: Session Information R-environment",
  table.attr = "style = \"color:black;\"") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),full_width = F, font_size = 12, position = "left", html_font = "Times New Roman") %>%
  scroll_box(width = "100%", height = "200px")
}
```

```{r Reproducibility-SessionInfo-R-MANY-packages-lanscape, echo=FALSE, fig.align="left", message=FALSE, warning=FALSE, out.width='100%', results='asis'}

if (Knit_to_Word) {
  
  } else {
df_session_packages <- devtools::session_info()$packages %>% 
  as.data.frame(.) %>% 
  filter(attached == TRUE) %>% 
  dplyr::select(loadedversion, date) %>% 
  rownames_to_column

colnames(df_session_packages) <- c("Package", "Loaded version", "Date")

df_session_packages %>% 
  kable(align = "c", table.attr = "style = \"color:black;\"", 
        caption = "Reproducibility: Package Information") %>% #center the column  
  kable_styling(bootstrap_options = c("striped", "hover"),full_width = F, font_size = 12, position = "left", html_font = "Times New Roman") %>%
  scroll_box(width = "100%", height = "200px")
}
```


Small change to test symlink and git push