---
title: "Bland Altman analysis of two methods Notebook"
output: html_notebook
---

This documents the code for performing a Bland-Altman analysis of two methods, and preparing graphs illustrating the results. It relies on the blandr package, as well as tidyverse and ggplot2
Created KPI Feb 20, 2020


```{r}
library (tidyverse)
library (blandr)

# Create dataframe and extract specific data for Bland Altman analysis

MT <- read.csv("/Volumes/InmanDrive4/Dropbox (Personal)/KPI/University_of_Dundee/Dissertation/Data/Software/LRmixStudio-2/LRMix_scripts/MT_LRM_gui_script.csv")

# MT$TotDNA <- as.factor(MT$TotDNA) # Need to turn the TotDNA into a factor (rather than integer) in order to plot using it as a group in ggplot (integer can't be used to group values, unless binning them). Before writing this step, I got an error in later code that MT$DNA did not exist, which of course it did. Finally realized you can't group by integers. 
```


# Subset dataframe with LR from one program (including columns needed for the subsetting task)
```{r}

baG <- MT %>%
  
  
  select(SampleName,TotlogLR,Source) %>%
  filter(Source == "gui")

# baOff <- na.omit(baOff)

# Subset dataframe with LR from the other program (including columns needed for the subsetting task)
baS <- MT %>%
  
  select(SampleName,TotlogLR,Source) %>%
  filter(Source == "script")
#baOn <- na.omit(baOn)
```

# Merge subsetted dataframes into one for Bland Altman calculations
```{r}

ba <- baG %>%
  
  left_join(baS, by = "SampleName", suffix = c(".gui", ".scr")) %>% 
  #inner_join(baOn, by = NULL)# %>%
  select(SampleName,TotlogLR.gui,TotlogLR.scr)

#ba <- filter(ba, TCon.Off == TCon.On)

# ba <- ba[-25,]

```

# Check on data to ensure integrity and adherence to proper format
```{r}

blandr.data.preparation(ba$TotlogLR.gui, ba$TotlogLR.scr, sig.level = 0.95) 


#Prepares statistics used in calculating the Bland-Altman plots for methods Lab Retriever and likeLTD)
bland.stats <- blandr.statistics(ba$TotlogLR.gui, ba$TotlogLR.scr) 
```

# Standard bland altman plot, using the script from the blandr package
```{r}

blandr.draw(ba$TotlogLR.gui, ba$TotlogLR.scr, method1name = "EFM gui LR",
            method2name = "EFM script LR",
            plotTitle = "Bland-Altman plot: LRmix gui vs. script",
            sig.level = 0.95, LoA.mode = 1, annotate = FALSE, ciDisplay = TRUE,
            ciShading = TRUE, normalLow = FALSE, normalHigh = FALSE,
            lowest_y_axis = FALSE, highest_y_axis = FALSE, point_size = 0.8,
            overlapping = FALSE, plotter = "ggplot2", x.plot.mode = "means",
            y.plot.mode = "difference", plotProportionalBias = FALSE,
            plotProportionalBias.se = TRUE, assume.differences.are.normal = TRUE)
```


# Various plot options using ggplot; can use in addition to or separate from the blandr.draw function

# Scatter plot with 4 plots, one for each amount of Total DNA (TotDNA) of the sample
```{r}

ggplot(ba, aes(x = bland.stats$means, y = bland.stats$differences)) +
  geom_point() + geom_text(aes(label=Sample), hjust=2, color = "blue") +
  facet_wrap(~ TCon.On) +
  labs(title = "Bland-Altman plot \nlikeLTD stutter Off/On", x = "Means", y = "Differences")

ggsave("BAplot_facet_pts.pdf", width = 13, height = 10)
 #+
  # theme_bw()

```


# Scatter plot with single plot, each amount of TotDNA color-coded
```{r}

ggplot(data = ba) +
  geom_point(mapping = aes(x = bland.stats$means, y = bland.stats$differences, color = TotDNA.On)) +
  labs(title = "Bland-Altman plot \n likeLTD stutter On", x = "Means", y = "Differences") +
  theme_bw()

```


# Smoothed plots of 4 DNA amounts (not that informative)
```{r}

ggplot(data = ba) +
  geom_smooth(mapping = aes(x = bland.stats$means, y = bland.stats$differences, color = TotDNA.On),
              show.legend = TRUE) +
  labs(title = "Bland-Altman plot \nlikeLTD stutter off/on", x = "Means", y = "Differences") +
  theme_bw()

```


# All data with a single smoothed line, with TotDNA color-coded
```{r}

ggplot(data = ba, mapping = aes(x = bland.stats$means, y = bland.stats$differences)) +
  geom_point(mapping = aes(color = TotlogLR.gui, shape = TotlogLR.gui)) +
  geom_smooth() +
  labs(title = "Bland-Altman plot \nlikeLTD stutter Off/On", x = "Means", y = "Differences") +
  theme_bw()

```


# All data with points plotted and smoothed line for each amt of TotDNA
```{r}

ggplot(data = ba, mapping = aes(x = bland.stats$means, y = bland.stats$differences)) +
  geom_point(mapping = aes(color = TotDNA.On, shape = TotDNA.On)) +
  geom_smooth(mapping = aes(x = bland.stats$means, y = bland.stats$differences, color = TotDNA.On),
            show.legend = TRUE) +
            labs(title = "Bland-Altman plot \nlikeLTD stutter Off/On", x = "Means", y = "Differences") +
            theme_bw()

```


# 4 plots, one for each amt of TotDNA, including smooth line, color, and shapes
```{r}

ggplot(data = ba, mapping = aes(x = bland.stats$means, y = bland.stats$differences)) +
      geom_point(mapping = aes(color = TotDNA.On, shape = TotDNA.On)) +
      geom_smooth(mapping = aes(x = bland.stats$means, y = bland.stats$differences, color = TotDNA.On),
              show.legend = TRUE) +
      facet_wrap(~ TCon.On) +
      labs(title = "Bland-Altman plot \nlikeLTD stutter Off/On", x = "Means", y = "Differences") +
      theme_bw()
```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

