---
title: "Bland Altman analysis of two methods Notebook"
output:
  html_document:
    df_print: paged
---

This documents the code for performing a Bland-Altman analysis of two methods, and preparing graphs illustrating the results. It relies on the blandr package, as well as tidyverse and ggplot2 Created KPI Feb 20, 2020

```{r}
library (tidyverse)
library (blandr)
library(ggthemes)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(readr)
library(dplyr)
library(kableExtra)
library(knitr)
library(ggpubr)
library(janitor)

source("/Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/Lab Notebooks/library.R")

```

```{r}

# Create dataframe and extract specific data for Bland Altman analysis

MT <- read.csv("/Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/Data/Output_files/AllResults_Data/AllResults_Table.csv")

MT$TotDNA <- as.factor(MT$TotDNA) # Need to turn the TotDNA into a factor (rather than integer) in order to plot using it as a group in ggplot (integer can't be used to group values, unless binning them). Before writing this step, I got an error in later code that MT$DNA did not exist, which of course it did. Finally realized you can't group by integers. 
```

# Subset dataframe with LR from EFM with stutter present (including columns needed for the subsetting task)

```{r}

baEFMoff <- MT %>%
  select(Sample,Program,TotlogLR,Eff,StFilt,TCon,TotDNA) %>%
  filter(Program == "EFM",StFilt =="off")

baEFMoff <- na.omit(baEFMoff)

# Subset dataframe with LR from EFM with stutter absent (including columns needed for the subsetting task)
baEFMon <- MT %>%
  select(Sample,Program,TotlogLR,Eff,StFilt,TCon,TotDNA) %>%
  filter(Program == "EFM",StFilt == "on")
baEFMon <- na.omit(baEFMon)
```

# Merge subsetted dataframes into one for Bland Altman calculations

```{r}

baEff <- baEFM %>%
  inner_join(baLTD, by = c("Sample","TCon")) %>% 
  select(Sample,Eff.x,Eff.y,TotDNA.x,TotDNA.y)


```

# Check on data to ensure integrity and adherence to proper format

```{r}

blandr.data.preparation(baEff$Eff.x, baEff$Eff.y, sig.level = 0.95) 

#Prepares statistics used in calculating the Bland-Altman plots for methods Lab Retriever and likeLTD)
bland.statsEff <- blandr.statistics(baEff$Eff.x, baEff$Eff.y) 
```

# Standard bland altman plot, using the script from the blandr package

```{r}

blandr.draw(baEff$Eff.x, baEff$Eff.y, method1name = "EFM",
            method2name = "likeLTD",
            plotTitle = "Bland-Altman plot:\nEFM/likeLTD Efficiency: stutter absent",
            sig.level = 0.95, LoA.mode = 1, annotate = FALSE, ciDisplay = TRUE,
            ciShading = TRUE, normalLow = FALSE, normalHigh = FALSE,
            lowest_y_axis = FALSE, highest_y_axis = FALSE, point_size = 0.8,
            overlapping = FALSE, plotter = "ggplot2", x.plot.mode = "means",
            y.plot.mode = "difference", plotProportionalBias = FALSE,
            plotProportionalBias.se = TRUE, assume.differences.are.normal = TRUE)

#ggsave2("/Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/Lab Notebooks/Statistical Testing and EDA templates/Bland_Altman/BA_EFMvlikeLTD.pdf", width = 7, height = 5)
```


```{r}

balLR <- baEFM %>%
  inner_join(baLTD, by = c("Sample","TCon")) %>% 
  select(Sample,TotlogLR.x,TotlogLR.y,TotDNA.x,TotDNA.y)


```

# Check on data to ensure integrity and adherence to proper format

```{r}

blandr.data.preparation(balLR$TotlogLR.x, balLR$TotlogLR.y, sig.level = 0.95) 

#Prepares statistics used in calculating the Bland-Altman plots for methods Lab Retriever and likeLTD)
bland.statslLR <- blandr.statistics(balLR$TotlogLR.x, balLR$TotlogLR.y) 

```

# Standard bland altman plot, using the script from the blandr package

```{r}

blandr.draw(balLR$TotlogLR.x, balLR$TotlogLR.y, method1name = "EFM",
            method2name = "likeLTD",
            plotTitle = "Bland-Altman plot:\nEFM/likeLTD Total logLR: stutter absent",
            sig.level = 0.95, LoA.mode = 1, annotate = FALSE, ciDisplay = TRUE,
            ciShading = TRUE, normalLow = FALSE, normalHigh = FALSE,
            lowest_y_axis = FALSE, highest_y_axis = FALSE, point_size = 0.8,
            overlapping = FALSE, plotter = "ggplot2", x.plot.mode = "means",
            y.plot.mode = "difference", plotProportionalBias = FALSE,
            plotProportionalBias.se = TRUE, assume.differences.are.normal = TRUE)

#ggsave2("/Volumes/InmanDrive4/Dropbox/Dropbox/KPI/University_of_Dundee/Dissertation/Lab Notebooks/Statistical Testing and EDA templates/Bland_Altman/BA_EFMvlikeLTD.pdf", width = 7, height = 5)
```
# Various plot options using ggplot; can use in addition to or separate from the blandr.draw function

# Scatter plot with 4 plots, one for each amount of Total DNA (TotDNA) of the sample

```{r}

ggplot(balLR, aes(x = bland.statslLR$means, y = bland.statslLR$differences)) +
  geom_point() + geom_text(aes(label=Sample), hjust=2, color = "blue") +
  facet_wrap(~ TotDNA.y) +
  labs(title = "Bland-Altman plot \nEFM/likeLTD stutter absent", x = "Means", y = "Differences")

#ggsave("BAplot_facet_pts.pdf", width = 13, height = 10)
 #+
  # theme_bw()

```

# Scatter plot with single plot, each amount of TotDNA color-coded

```{r}

ggplot(data = balLR) +
  geom_point(mapping = aes(x = bland.statslLR$means, y = bland.statslLR$differences, color = TotDNA.y)) +
  labs(title = "Bland-Altman plot \nEFM/likeLTD stutter absent", x = "Means", y = "Differences") +
  theme_bw()

```

# Smoothed plots of 4 DNA amounts (not that informative)

```{r}

ggplot(data = balLR) +
  geom_smooth(mapping = aes(x = bland.statslLR$means, y = bland.statslLR$differences, color = TotDNA.y),
              show.legend = TRUE) +
  labs(title = "Bland-Altman plot \nEFM/likeLTD stutter absent", x = "Means", y = "Differences") +
  theme_bw()

```

# All data with a single smoothed line, with TotDNA color-coded

```{r}

ggplot(data = balLR, mapping = aes(x = bland.statslLR$means, y = bland.statslLR$differences)) +
  geom_point(mapping = aes(color = TotDNA.y, shape = TotDNA.y)) +
  geom_smooth() +
  labs(title = "Bland-Altman plot \nEFM/likeLTD stutter absent", x = "Means", y = "Differences") +
  theme_bw()

```

# All data with points plotted and smoothed line for each amt of TotDNA

```{r}

ggplot(data = balLR, mapping = aes(x = bland.statslLR$means, y = bland.statslLR$differences)) +
  geom_point(mapping = aes(color = TotDNA.y, shape = TotDNA.y)) +
  geom_smooth(mapping = aes(x = bland.statslLR$means, y = bland.statslLR$differences, color = TotDNA.y),
            show.legend = TRUE) +
            labs(title = "Bland-Altman plot \nEFM/likeLTD stutter off", x = "Means", y = "Differences") +
            theme_bw()

```

# 4 plots, one for each amt of TotDNA, including smooth line, color, and shapes

```{r}

ggplot(data = balLR, mapping = aes(x = bland.statslLR$means, y = bland.statslLR$differences)) +
      geom_point(mapping = aes(color = TotDNA.y, shape = TotDNA.y)) +
      geom_smooth(mapping = aes(x = bland.statslLR$means, y = bland.statslLR$differences, color = TotDNA.y),
              show.legend = TRUE) +
      facet_wrap(~ TotDNA.y) +
      labs(title = "Bland-Altman plot \nEFM/likeLTD stutter off", x = "Means", y = "Differences") +
      theme_bw()
```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
